\section{The Most Common Mistake at Each Sprint}

{\centering\small
\begin{longtable}{c >{\raggedright\arraybackslash}p{4.5cm} >{\raggedright\arraybackslash}p{5.8cm}}
\toprule
\textbf{Sprint} & \textbf{Most Common Mistake} & \textbf{Prevention} \\
\midrule
\endhead
0 & Starting to design before understanding the problem & Spend Sprint~0 listening, not solving. Ask questions, take notes. \\
1 & SOW with no ``Out of Scope'' section & Explicitly list what you will NOT do. Get client agreement. \\
2 & Generating only one concept & Use brainstorming, SCAMPER, morphological chart. Generate $\geq$15 ideas. \\
3 & Decision matrix rigged to support a predetermined choice & Add sensitivity analysis. Vary weights by $\pm$20\%. \\
4 & PoC that does not retire the key technical risk & Identify the risk first, then design the PoC to test it specifically. \\
5 & Skipping the Mock PDR & Schedule the mock 3 days before. Fresh eyes catch problems you miss. \\
6 & Not ordering long-lead components & Order PCBs and specialty parts NOW. They arrive in time for SDII. \\
\midrule
7 & Skipping the Hazard Assessment & Required before fabrication. PA must approve. No shortcuts. \\
8 & BOM missing fasteners and consumables & Walk through the CAD assembly. Every visible component needs a BOM line. \\
9--10 & Big-bang integration & Test each subsystem individually before connecting one at a time. \\
11 & Leaving all testing to Sprint~12 & Start V\&V testing NOW. Every week of margin is one redesign cycle. \\
12--13 & Documentation debt & Write report sections during the sprint the work is done, not at the end. \\
14 & Incomplete VCRM at FDR & Every requirement must be dispositioned: Verified, Failed (with disposition), Waived, or Deferred. \\
\bottomrule
\end{longtable}
}


% ═══════════════════════════════════════
