\section{SDI Sprint 6: Design Refinement, Next Steps, SDII Handoff}
\textit{Weeks 14--15 $|$ Course Text: Chapters 8, 10}

\begin{priorities}
\begin{enumerate}[nosep]
\item Incorporate PDR feedback: categorize, assign owners, complete.
\item Baseline the SDRD (freeze; all future changes under change control; see Course Text Ch.~11 for the change control process).
\item Begin procurement of long-lead-time components for SDII.
\item Prepare SDII transition documentation.
\item Send Next Steps Email to client.
\end{enumerate}
\end{priorities}

\begin{protip}
Order long-lead-time components \emph{now}. Do not wait until SDII Sprint~8. Custom PCBs take 2--4 weeks. Specialty sensors take 2--4 weeks. Every week you wait in Sprint~6 is a week less margin in SDII.
\end{protip}

\begin{warnbox}
\textbf{Competition Track:} If your competition is in spring semester, start travel planning NOW. Mines travel approvals require a minimum 8-week lead time. Late travel requests create administrative problems and may not be approved.
\end{warnbox}

\subsection{Deliverables}
\begin{itemize}[nosep,leftmargin=1.5em]
\item PDRt Update (incorporating feedback)
\item Next Steps Email to client
\item Peer Evaluation \#2
\item End-of-Semester Course Survey
\item Sprint 5 Review / Retrospective / Sprint 6 documentation
\end{itemize}

\subsection{Preparing for SDII}
The semester break between SDI and SDII is \emph{not} dead time. Before leaving for winter break: ensure all long-lead-time orders are placed, finalize your SDII transition documentation, and confirm your client's availability for the first week of SDII (Sprint~7). Week numbering resets to~1 in SDII. Review the Sprint~7 section of this guide before the first day of classes so you are ready to hit the ground running.

% ═══════════════════════════════════════
