\section{SDI Sprint 0: Project Expo, Team Formation, Client Kickoff}
\textit{Weeks 1--2 $|$ Course Text: Chapters 1, 2}

\begin{priorities}
\begin{enumerate}[nosep]
\item Attend the Project Expo and complete the Project Selection Survey.
\item Meet your team, PA, and client.
\item Prepare for and conduct the Client Kickoff Meeting.
\item Set up the team repository (Teams group, file structure).
\item Begin the Individual Design Log (a personal notebook---physical or digital---where you record design decisions, sketches, calculations, and meeting notes throughout both semesters).
\end{enumerate}
\end{priorities}

\begin{faqbox}[What should I do before the kickoff meeting?]
Research the client's organization and industry. Prepare a slide deck for project discovery (you will present these during the Sprint~1 client kickoff). Write 10 open-ended questions. Assign roles: facilitator, scribe, timekeeper. See Course Text Ch.~2 for the full kickoff workflow.
\end{faqbox}

\begin{warnbox}
Do not start designing during Sprint~0. Your job this sprint is to \emph{understand the problem}, not to solve it. Listen more than you talk at the kickoff meeting. The first idea is rarely the best idea.
\end{warnbox}

\begin{protip}
Send a follow-up email to the client within 24 hours of the kickoff meeting summarizing key decisions, open questions, and next steps. Copy your PA. This establishes your professionalism from Day~1.
\end{protip}

\begin{faqbox}[Competition Track: What to do in Sprint 0]
\begin{itemize}[nosep]
\item Read the \emph{entire} competition rulebook before the kickoff meeting.
\item Create a preliminary compliance matrix mapping competition rules to design requirements.
\item Identify the competition registration deadline and add it to your project timeline.
\item Determine if the competition is Build Track or Paper Track (or Hybrid).
\end{itemize}
\end{faqbox}

\subsection{Deliverables}
\begin{itemize}[nosep,leftmargin=1.5em]
\item What to Expect from SD Quiz (Canvas)
\item Project Selection Survey (by 11:30 AM on the designated date)
\item Prof.~Dev.~-- IP Module, Licensure Module
\item Email to client to introduce team with team slide
\end{itemize}


% ═══════════════════════════════════════
