\section{The Capstone Sprint: A Typical Week}

\begin{warnbox}
\textbf{This course meets formally on Tuesdays and Thursdays only} (lectures and recitations). The Monday, Wednesday, and Friday activities shown below are \emph{self-directed} team work sessions that you are responsible for organizing outside of class time. Capstone requires significant independent effort beyond scheduled meetings. Teams that rely solely on Tuesday/Thursday class time will fall behind.
\end{warnbox}

\begin{warnbox}
\textbf{Attendance at every lecture and recitation is mandatory.} Key deliverable instructions, client feedback, design review logistics, and PA assessments occur during these sessions. Material covered in lecture and recitation cannot always be recovered from slides alone. Missing a session means missing information your team needs to stay on track.
\end{warnbox}

Your team's weekly rhythm should follow this general pattern (adapt to your specific schedule):

\begin{center}\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Day} & \textbf{Activities} \\
\midrule
Monday & \textit{Self-directed}: Team stand-up (15 min). Update task board. Identify blockers. \\
Tuesday & \textbf{Scheduled}: Lecture / Recitation / Lab session. Design, analysis, or fabrication work. \\
Wednesday & \textit{Self-directed}: Work session. PA check-in (show progress, surface issues). \\
Thursday & \textbf{Scheduled}: Lecture / Recitation / Lab session. Send client status email (biweekly). \\
Friday & \textit{Self-directed}: Work session. Documentation. End of Week 2: Sprint Review + Retrospective. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{protip}
The three non-negotiable weekly touchpoints: (1)~the stand-up (team coordination), (2)~the PA check-in (accountability), and (3)~the client communication (professionalism). Everything else flexes around your project needs. Plan for 6--10 hours per week of team work beyond scheduled class time.
\end{protip}


% ═══════════════════════════════════════
