\chapter*{How to Use This Document}
\addcontentsline{toc}{chapter}{How to Use This Document}

This reference document accompanies the EDNS~491 and EDNS~492 course syllabi and the Canvas course site. It is organized to support you across the entire two-semester capstone design experience, from your first client meeting through the Design Showcase poster session.

\textbf{Three documents work together:}
\begin{enumerate}[nosep,leftmargin=1.5em]
\item \textbf{This Course Text}: deep technical and process content organized in five parts. Consult it when you need the methodology, design procedures, project management frameworks, communication guidance, technical reference, or safety protocols. Start with the ``What Should I Be Reading Right Now?'' triage guide on the next page.
\item \textbf{The Course Syllabus}: assignment details, grading weights, policies, CLOs, and faculty contact information. The syllabus is the authoritative source for deadlines and grading.
\item \textbf{Canvas}: assignment rubrics, submission portals, scheduling, lecture materials, professional development modules, purchasing forms, and the Labriola InnoHub enrollment site.
\end{enumerate}

\textbf{Part~I: The Design Process} (Chapters~1--6) covers the complete engineering design lifecycle from problem definition through verification and validation. These chapters support your work in chronological order as your project progresses through design reviews.

\textbf{Part~II: Major Project Milestones} (Chapters~7--10) provides detailed guidance for the Statement of Work, PDR, CDR, and FDR, with deliverable checklists, evaluation criteria, and common failure patterns at each gate.

\textbf{Part~III: Project Management and Operations} (Chapters~11--16) covers hybrid project management, fabrication spaces and resources, team dynamics, budgeting, purchasing procedures, and travel planning. These chapters are practical operations guides that you will reference repeatedly throughout both semesters.

\textbf{Part~IV: Professional Skills} (Chapters~17--20) covers client engagement, written and oral communication, and career preparation. These skills distinguish capstone from prior coursework and directly affect your individual performance evaluations.

\textbf{Part~V: Safety, Compliance, and Professional Practice} (Chapters~21--22) covers EHS requirements, professional ethics, PE licensure, and course policies.

\textbf{Appendices} include a glossary, quick reference tables, ``Why Capstone Projects Fail'' (12 common patterns), checklists, and references. ABET outcome mapping and the NCEES Rules of Professional Conduct are covered in Chapter~22.

\subsection*{Reading by Role}
\begin{itemize}[nosep,leftmargin=1.5em]
\item \textbf{Scrum Master / Project Manager}: Chapters~11 (PM), 13 (Team Dynamics), 7 (SOW), 8--10 (Milestones)
\item \textbf{Chief Financial Officer}: Chapters~14 (Budgeting), 15 (Purchasing), 16 (Travel)
\item \textbf{Technical Lead}: Chapters~3 (Requirements), 4 (Concept Gen), 6 (V\&V), 9 (CDR)
\item \textbf{Communication Lead}: Chapters~17 (Client), 18 (Written), 19 (Oral), 13 (Team)
\item \textbf{Fabrication Lead}: Chapter~12 (Fab Spaces), 21 (EHS), Front Matter (Safety)
\end{itemize}

\subsection*{Visual Guide: Box Types Used in This Document}
Throughout this document, colored boxes convey different types of information. Scan for these when reviewing a chapter:

\begin{center}\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Box} & \textbf{Purpose} \\
\midrule
\textcolor{MinesBlue}{\textbf{Key Concept}} & Core definitions and frameworks (the ``must know'' material). \\
\textcolor{WarnOrange}{\textbf{Warning / Common Pitfall}} & Frequent mistakes and failure patterns that cost teams time, money, or grades. Read these carefully. \\
\textcolor{AccentTeal}{\textbf{Tip}} & Best practices and professional advice. Not required, but will save you effort. \\
\textcolor{AccentTeal}{\textbf{Example}} & Worked examples and case studies illustrating a concept in practice. \\
\textcolor{GreenShield}{\textbf{Case Study}} & Extended real-world applications, typically project-specific. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection*{The First 30 Days: A Role-Based Checklist}
Use these checklists to ensure each role is on track by the end of Sprint~2:

\textbf{Scrum Master / Project Manager}: \\
$\square$ Task board created and populated with Sprint~0/1 items \\
$\square$ Weekly PA check-in meeting scheduled \\
$\square$ First Team Retrospective conducted \\
$\square$ Meeting norms established (agenda template, scribe rotation, minutes distribution)

\textbf{Chief Financial Officer}: \\
$\square$ Introduced yourself to Capstone Purchasing (CP) \\
$\square$ Reviewed the approved vendor list \\
$\square$ Created and shared the initial Budget Tracker spreadsheet \\
$\square$ Downloaded current tax-exempt certificate from Canvas

\textbf{Technical Lead}: \\
$\square$ Requirements elicitation started (client questions mapped to SDRD structure) \\
$\square$ Prior art and competitive benchmarking initiated \\
$\square$ Engineering standards identified for the project domain

\textbf{Communication Lead}: \\
$\square$ Client communication cadence established (frequency, format, who is CC'd) \\
$\square$ Kickoff follow-up email sent within 24 hours \\
$\square$ File naming and version control conventions defined for the team

\textbf{Fabrication Lead}: \\
$\square$ InnoHub equipment training registered (or completed) \\
$\square$ Available fabrication resources inventoried (shop capabilities, 3D printers, laser cutters) \\
$\square$ Safety training modules completed on Labriola Canvas site

\subsection*{Project Tracks}
This document supports all three project tracks. Where guidance differs by track, look for \textbf{Build Track}, \textbf{Paper Track}, and \textbf{Competition Track} labels. All tracks are held to equivalent standards of engineering rigor. The Paper Track is not a reduced-effort alternative; it substitutes physical fabrication with analytical depth, simulation fidelity, and documentation completeness. The Competition Track does not reduce requirements; it reorganizes their timing around an external schedule.


% ========== TRIAGE TABLE ==========
