\chapter*{References and Further Resources}
\addcontentsline{toc}{chapter}{References and Further Resources}

\section*{Engineering Design Process and Systems Engineering}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em]
\item K.~Ulrich and S.~Eppinger, \emph{Product Design and Development}, 7th~ed., McGraw-Hill, 2020.
\item INCOSE, \emph{Systems Engineering Handbook}, 5th~ed., Wiley, 2023.
\item S.~Pugh, \emph{Total Design: Integrated Methods for Successful Product Engineering}, Addison-Wesley, 1991.
\item G.~Pahl, W.~Beitz, J.~Feldhusen, and K.-H.~Grote, \emph{Engineering Design: A Systematic Approach}, 3rd~ed., Springer, 2007.
\end{enumerate}

\section*{Project Management}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item D.\,L.~Goetsch, \emph{Project Management for Engineering and Technology}, Pearson, 2015.
\item K.~Schwaber and J.~Sutherland, \emph{The Scrum Guide: The Definitive Guide to Scrum}, Scrum.org, Nov.\ 2020. [Online]. Available: \url{https://scrumguides.org/}
\item M.~Lacey, \emph{The Scrum Field Guide: Practical Advice for Your First Year}, 2nd~ed., Addison-Wesley, 2015.
\item Project Management Institute, \emph{A Guide to the Project Management Body of Knowledge (PMBOK Guide)}, 7th~ed., PMI, 2021.
\end{enumerate}

\section*{Client Engagement and Requirements}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item R.~Fitzpatrick, \emph{The Mom Test: How to Talk to Customers and Learn If Your Business Is a Good Idea When Everyone Is Lying to You}, CreateSpace, 2013.
\item D.~Clegg and R.~Barker, \emph{Case Method Fast-Track: A RAD Approach}, Addison-Wesley, 1994. (Origin of MoSCoW prioritization.)
\end{enumerate}

\section*{Team Dynamics}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item B.\,W.~Tuckman, ``Developmental Sequence in Small Groups,'' \emph{Psychological Bulletin}, vol.~63, no.~6, pp.~384--399, 1965.
\item B.\,W.~Tuckman and M.\,A.\,C.~Jensen, ``Stages of Small-Group Development Revisited,'' \emph{Group \& Organization Studies}, vol.~2, no.~4, pp.~419--427, 1977.
\end{enumerate}

\section*{Technical References}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item P.~Horowitz and W.~Hill, \emph{The Art of Electronics}, 3rd~ed., Cambridge University Press, 2015.
\item R.\,G.~Budynas and J.\,K.~Nisbett, \emph{Shigley's Mechanical Engineering Design}, 11th~ed., McGraw-Hill, 2020.
\end{enumerate}

\section*{Risk Analysis and Quality}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item SAE International, ``Potential Failure Mode and Effects Analysis in Design (Design FMEA) and Potential Failure Mode and Effects Analysis in Manufacturing and Assembly Processes (Process FMEA),'' SAE~J1739, Jan.\ 2009.
\item IEC, ``Failure Modes and Effects Analysis (FMEA and FMECA),'' IEC~60812, Ed.~3.0, Aug.\ 2018.
\end{enumerate}

\section*{Sustainability and Ethics}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item U.S.\ Green Building Council, ``LEED v4.1,'' 2023. [Online]. Available: \url{https://www.usgbc.org/leed}
\item Institute for Sustainable Infrastructure, ``Envision: Sustainable Infrastructure Framework,'' v3, 2018. [Online]. Available: \url{https://sustainableinfrastructure.org/envision/}
\item NSPE, \emph{Code of Ethics for Engineers}, National Society of Professional Engineers, rev.\ 2007.
\item NCEES, \emph{Model Rules}, Section~240: Rules of Professional Conduct. [Online]. Available: \url{https://ncees.org/licensure/laws-and-rules/}
\end{enumerate}

\section*{Safety Standards}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item IEC, ``Safety Requirements for Electrical Equipment for Measurement, Control, and Laboratory Use,'' IEC~61010-1, Ed.~3.1, 2020.
\item OSHA, ``General Duty Clause,'' Section 5(a)(1) of the Occupational Safety and Health Act of 1970, 29~U.S.C.\ \S654.
\item NFPA, ``NFPA 70E: Standard for Electrical Safety in the Workplace,'' National Fire Protection Association, 2024.
\end{enumerate}

\section*{Professional Communication}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item IEEE, ``IEEE Editorial Style Manual for Authors,'' IEEE Periodicals, 2024. [Online]. Available: \url{https://journals.ieeeauthorcenter.ieee.org/}
\end{enumerate}

\section*{Accreditation}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item ABET Engineering Accreditation Commission, \emph{Criteria for Accrediting Engineering Programs}, 2024--2025. [Online]. Available: \url{https://www.abet.org/accreditation/accreditation-criteria/}
\end{enumerate}

\section*{Design Methods and Ideation}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item A.\,F.~Osborn, \emph{Applied Imagination: Principles and Procedures of Creative Problem-Solving}, 3rd~ed., Scribner, 1963.
\item B.~Eberle, \emph{SCAMPER: Creative Games and Activities for Imagination Development}, Prufrock Press, 1996. (Originally published 1971.)
\item F.~Zwicky, \emph{Discovery, Invention, Research Through the Morphological Approach}, Macmillan, 1969.
\item J.\,M.~Benyus, \emph{Biomimicry: Innovation Inspired by Nature}, William Morrow, 1997.
\item J.~Elkington, \emph{Cannibals with Forks: The Triple Bottom Line of 21st Century Business}, Capstone Publishing, 1997.
\item W.~McDonough and M.~Braungart, \emph{Cradle to Cradle: Remaking the Way We Make Things}, North Point Press, 2002.
\end{enumerate}

\section*{Requirements and Systems Engineering Standards}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item IEEE, ``IEEE Guide for Information Technology---System Definition---Concept of Operations (ConOps) Document,'' IEEE Std~1362-1998, 1998.
\item ISO/IEC/IEEE, ``Systems and Software Engineering---Life Cycle Processes---Requirements Engineering,'' ISO/IEC/IEEE~29148:2018, 2018.
\item IEEE, ``IEEE Standard for System, Software, and Hardware Verification and Validation,'' IEEE Std~1012-2016, 2016.
\item NASA, \emph{NASA Systems Engineering Handbook}, Rev.~2, NASA/SP-2016-6105, 2016.
\end{enumerate}

\section*{Cost Estimation and Metrics}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item B.\,W.~Boehm, \emph{Software Engineering Economics}, Prentice-Hall, 1981.
\item S.~McConnell, \emph{Software Estimation: Demystifying the Black Art}, Microsoft Press, 2006.
\item G.\,T.~Doran, ``There's a S.M.A.R.T. Way to Write Management's Goals and Objectives,'' \emph{Management Review}, vol.~70, no.~11, pp.~35--36, 1981.
\end{enumerate}

\section*{Team Dynamics and Communication}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item A.~Edmondson, ``Psychological Safety and Learning Behavior in Work Teams,'' \emph{Administrative Science Quarterly}, vol.~44, no.~2, pp.~350--383, 1999.
\item E.~Derby and D.~Larsen, \emph{Agile Retrospectives: Making Good Teams Great}, Pragmatic Bookshelf, 2006.
\item A.\,W.\,B.~Brooks, ``Get Excited: Reappraising Pre-Performance Anxiety as Excitement,'' \emph{J. Experimental Psychology: General}, vol.~143, no.~3, pp.~1144--1158, 2014.
\end{enumerate}

\section*{Safety and Regulatory}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em,resume]
\item IEC, ``Effects of Current on Human Beings and Livestock,'' IEC~60479-1, Ed.~5.0, 2018.
\item OSHA, ``Occupational Noise Exposure,'' 29~CFR~1910.95.
\item European Parliament, ``Directive 2011/65/EU on the Restriction of the Use of Certain Hazardous Substances in Electrical and Electronic Equipment (RoHS~2),'' 2011.
\item S.~Van~Wassenbergh, E.\,J.~Ortlieb, A.~Mielke, C.~B\"ohmer, R.\,E.~Shadwick, and A.~Abourachid, ``Woodpeckers Minimize Cranial Absorption of Shocks,'' \emph{Current Biology}, vol.~32, no.~14, pp.~3188--3194, Jul.\ 2022.
\end{enumerate}

\section*{Mines Resources}
\begin{itemize}[leftmargin=1.5em]
\item \textbf{EHS}: \url{https://www.mines.edu/ehs/}
\item \textbf{Labriola Innovation Hub}: \url{https://labriola.mines.edu/}
\item \textbf{Writing Center}: \url{https://writing.mines.edu/}
\item \textbf{Mines Online}: \url{https://online.mines.edu/}
\item \textbf{Capstone Purchasing}: capstonepurchasing@mines.edu
\item \textbf{NCEES FE Reference Handbook}: \url{https://ncees.org/exams/study-materials/}
\item \textbf{AskNature (biomimicry database)}: \url{https://asknature.org/}
\item \textbf{Google Patents}: \url{https://patents.google.com/}
\end{itemize}

