\chapter*{References and Further Resources}
\addcontentsline{toc}{chapter}{References and Further Resources}

\section*{Textbooks}
\begin{enumerate}[label={[\arabic*]},leftmargin=2em]
\item D.\,L.~Goetsch, \emph{Project Management for Engineering and Technology}, Pearson, 2015.
\item M.~Lacey, \emph{The Scrum Field Guide: Practical Advice for Your First Year}, Addison-Wesley, 2012.
\item K.~Ulrich and S.~Eppinger, \emph{Product Design and Development}, 7th~ed., McGraw-Hill, 2020.
\item P.~Horowitz and W.~Hill, \emph{The Art of Electronics}, 3rd~ed., Cambridge University Press, 2015.
\item R.\,G.~Budynas and J.\,K.~Nisbett, \emph{Shigley's Mechanical Engineering Design}, 11th~ed., McGraw-Hill, 2020.
\item INCOSE, \emph{Systems Engineering Handbook}, 5th~ed., Wiley, 2023.
\end{enumerate}

\section*{Standards}
\begin{itemize}[leftmargin=1.5em]
\item ABET Engineering Accreditation Commission, \emph{Criteria for Accrediting Engineering Programs}, 2024--2025.
\item NSPE, \emph{Code of Ethics for Engineers}, National Society of Professional Engineers.
\item NCEES, \emph{Model Rules}, Section~240: Rules of Professional Conduct.
\end{itemize}

\section*{Mines Resources}
\begin{itemize}[leftmargin=1.5em]
\item \textbf{EHS}: \url{https://www.mines.edu/ehs/}
\item \textbf{Labriola Innovation Hub}: \url{http://labriola.mines.edu/}
\item \textbf{Writing Center}: \url{https://writing.mines.edu/}
\item \textbf{Mines Online}: \url{https://online.mines.edu/}
\item \textbf{Capstone Purchasing}: capstonepurchasing@mines.edu
\end{itemize}

